\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2020-12-19 popl-2019-journal-v/paper.tex
Polish and shorten section 4 a bit more.

\lefttitle{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using \ldots}
\righttitle{Journal of Functional Programming}

@@

  \begin{subfigure}[h]{.8\linewidth}
    \begin{mathpar}
  \inferrule*[Right=E-Sealed-Code]{
    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
    \sigma \in \sigcloss
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
  }
  \and
  \inferrule*[Right=E-Word]{
    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
  }
    \end{mathpar}
    \caption{Export well-formedness}
    \label{fig:well-formed-export}
  \end{subfigure}
  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
  \label{fig:well-formed-aux}
\end{figure}
\figurename~\ref{fig:well-formed-word} defines the well-formedness judgement $\vdash_{\mathrm{comp-word}}$ for words in a component's data memory.
The \textsc{W-Data} rule says all data, i.e.\ integers, are well-formed words.
The \textsc{W-Capability} rule specifies that all well-formed capabilities in data memory must at most have read and write permission.
Further, if the capability is linear, then the range of authority must be within the linear address space owned by this address.
Finally, if it is non-linear, then it must be in the non-linear address space.
The \textsc{W-Sealed-Capability} allows capabilities sealed with a closure seal if the underlying capability is itself a well-formed word.
Because of the last requirement, the data memory cannot initially contain sealed code capabilities.
However, a component is free to place sealed data capabilities in memory during execution.

\figurename~\ref{fig:well-formed-export} defines the well-formedness judgement $\vdash_{\mathrm{comp-export}}$ for component exports.
Essentially, \textsc{E-Word} says that any well-formed word is also allowed as an export.
In particular, sealed capabilities for data memory are well-formed, which allows components to export the data part of a sealed capability pair.
Additionally, \textsc{E-Sealed-Code} allows read-execute capabilities to the component's code address space when they have been sealed with a closure seal.
Together, these two rules allow components to export closures as sealed code-data capability pairs.

Finally, \figurename~\ref{fig:well-formed-code} defines the well-formedness judgement $\vdash_{\mathrm{comp-code}}$ for component's code memory.
Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each word in isolation, because it depends on whether seals for calls are located in the place we expect them to be located.
Therefore, the main judgement considers the entire code memory $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ and is defined in terms of an auxiliary judgement for an individual memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
The $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$ judgement has two rules.
The \textsc{C-Seals} rule allows set-of-seals capabilities for closure and return seals.
The \textsc{C-Instr} rule allows integers, including encoded instructions.
However, if the code memory contains a call instruction in a trusted component (i.e.\ addresses in $\ta$), then there must be a set-of-seals capability available at the expected address, containing the return seal for this call.
To ensure that return seals are at most used by one call, the $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ judgement partitions all the available return seals over code addresses and \textsc{C-Instr} requires that the seal used by a call is in $\sigrets[\mathrm{owned}]$.
Further, the \textsc{C-Mem} judgement requires the set of return seals to be disjoint from the set of closure seals, so return seals cannot be used to seal non-return capabilities.
% domi: I have investigated this and I don't think we use this any more, as pointed out by Lau below.
% To rule out certain corner cases, we require the code memory to have no \textit{hidden calls}.
% \lau{Do we actually use the hidden calls in the proofs? I vaguely recall this was added to make sure that splicing would not create a call. However, we also have the padding which, at least in part, takes care of this.}
% \begin{definition}[No hidden calls]
%   \label{def:no-hidden-calls}
%   We say that a memory fragment $\mscode$ has no hidden calls iff
%   \[
%     \begin{array}{l}
%       \forall \aaddr \in \dom(\mscode) \ldotp \\
%       \quad \forall i \in [0,\calllen - 1] \\
%       \qquad \mscode(a + i) = \scall[i]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow \\
%       \qquad \quad (\dom(\mscode) \supseteq [\aaddr-i,\aaddr + \calllen - i - 1] \wedge\\
%       \qquad \qquad \mscode([\aaddr-i,\aaddr + \calllen - i -1]) =
%       \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \vee\\
%       \qquad \quad \exists j \in [\aaddr-i,\aaddr + \calllen - i -1] \cap \dom(\mscode) \ldotp \mscode(j) \neq  \scall[j-\aaddr-i]{\offpc,\offsigma}{r_1}{r_2}
%     \end{array}
%   \]
% \end{definition}
% The definition requires that if code memory contains an instruction that may be part of a call, then either
% \begin{itemize}
% \item it is a part of a call which falls entirely within the code memory or 
% \item there is some other instruction within the code memory, that witnesses the fact that it is not part of a call.
% \end{itemize}
% domi: update: no longer needed either, see TR.
% Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set-of-seals available.

@@

A well-formed component makes sure that a return seal is uniquely available to every call.
This is, however, not sufficient as it does not ensure that other parts of a program don't use the return seals.
We do not want to specify what non-call code should look like, so we just require it not to use the call seals.
This is what Item~\ref{item:return-seal-usage} ensures.
It says that if the configuration points to a seal-instruction, then either the instruction is part of a call and uses a return seal or the instruction seals part of a closure and uses a closure seal.
In order to construct a closure, one must seal a code capability and a data capability.
If this is an executable capability, then it should be reasonable as a pc.
On the other hand, if it is not an executable capability, then this must be the data part of the pair, so it should just be a reasonable word.
Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle.

>>>>

\lefttitle{\stktokens{}: Enforcing WBCF and LSE Using Linear Capabilities}
\righttitle{Journal of Functional Programming}

@@

  \begin{subfigure}[h]{.8\linewidth}
    \begin{mathpar}
  \inferrule*[Right=E-Sealed-Code]{
    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
    \sigma \in \sigcloss
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
  }
  \and
  \inferrule*[Right=E-Word]{
    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
  }{
    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
  }
    \end{mathpar}
    \caption{Export well-formedness}
    \label{fig:well-formed-export}
  \end{subfigure}
  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
  \label{fig:well-formed-aux}
\end{figure}
\figurename~\ref{fig:well-formed-word} defines the well-formedness judgement $\vdash_{\mathrm{comp-word}}$ for words in a component's data memory.
Well-formed rules are defined as integers, memory capabilities with at most read and write permission and well-formed words sealed with a closure seal.
The range of authority of linear and non-linear memory capabilities is required to be available in $A_{\mathrm{linear}}$ (where it is exclusively owned) and $A_{\mathrm{non-linear}}$, respectively.
Note that data memory is not allowed to initially contain sealed code capabilities.
However, a component is free to place sealed data capabilities in memory during execution.

\figurename~\ref{fig:well-formed-export} defines the well-formedness judgement $\vdash_{\mathrm{comp-export}}$ for component exports.
Essentially, well-formed exports are either well-formed words or sealed read-execute capabilities to the component's code memory, sealed with a closure seal.
Both cases together allow components to export closures as sealed code-data capability pairs.

Finally, \figurename~\ref{fig:well-formed-code} defines the well-formedness judgement $\vdash_{\mathrm{comp-code}}$ for components' code memory.
% Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each word in isolation, because it depends on whether seals for calls are located in the place we expect them to be located.
The judgement $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ is defined in terms of an auxiliary judgement for an individual memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
The rules allow either set-of-seals capabilities for closure and return seals, or integers, including encoded instructions.
If the code memory contains a call instruction in a trusted component (i.e.\ addresses in $\ta$), then there must be a set-of-seals capability available at the expected address, containing the return seal for this call.
Additionally, the judgement ensures that return seals are at most used by one call by partitioning
available return seals in $\sigrets[\mathrm{owned}]$ over code addresses for use in calls.
Further, return seals must be disjoint from closure seals.
% domi: I have investigated this and I don't think we use this any more, as pointed out by Lau below.
% To rule out certain corner cases, we require the code memory to have no \textit{hidden calls}.
% \lau{Do we actually use the hidden calls in the proofs? I vaguely recall this was added to make sure that splicing would not create a call. However, we also have the padding which, at least in part, takes care of this.}
% \begin{definition}[No hidden calls]
%   \label{def:no-hidden-calls}
%   We say that a memory fragment $\mscode$ has no hidden calls iff
%   \[
%     \begin{array}{l}
%       \forall \aaddr \in \dom(\mscode) \ldotp \\
%       \quad \forall i \in [0,\calllen - 1] \\
%       \qquad \mscode(a + i) = \scall[i]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow \\
%       \qquad \quad (\dom(\mscode) \supseteq [\aaddr-i,\aaddr + \calllen - i - 1] \wedge\\
%       \qquad \qquad \mscode([\aaddr-i,\aaddr + \calllen - i -1]) =
%       \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \vee\\
%       \qquad \quad \exists j \in [\aaddr-i,\aaddr + \calllen - i -1] \cap \dom(\mscode) \ldotp \mscode(j) \neq  \scall[j-\aaddr-i]{\offpc,\offsigma}{r_1}{r_2}
%     \end{array}
%   \]
% \end{definition}
% The definition requires that if code memory contains an instruction that may be part of a call, then either
% \begin{itemize}
% \item it is a part of a call which falls entirely within the code memory or 
% \item there is some other instruction within the code memory, that witnesses the fact that it is not part of a call.
% \end{itemize}
% domi: update: no longer needed either, see TR.
% Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set-of-seals available.

@@

A well-formed component makes sure that a return seal is uniquely available to every call.
This is, however, not sufficient as it does not ensure that other parts of a program don't use the return seals.
We do not want to specify what non-call code should look like, so we just require it not to use the call seals.
This is what Item~\ref{item:return-seal-usage} ensures.
It says that if the configuration points to a seal-instruction, then either the instruction is part of a call and uses a return seal or the instruction seals part of a closure and uses a closure seal.
In the latter case, the sealed capability must be reasonable as a pc if it is executable or just a reasonable word if it isn't.
Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle.
